---
title: "Team 5 - Project 3"
author: "Ariba Mandavia, Jose Fuentes, Marco Castro, Steven Gonzalez"
date: "2024-10-14"
output: html_document
---

# Using DuckDB
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(dbplyr)
library(tidyr)
library(readr)
library(stringr)
library(arrow)
library(duckdb)

```

## Initializing DuckDB

This bit just initializes DuckDB. First create a connection to a new DuckDB then register your files. _duckdb_register_ basically just converts your file into a database table.

Note: I may try to make a more programatic call later on. 

```{r inti_duck_db}
# initialize a new instance of DuckDB 
# and create a connection
con <- dbConnect(duckdb::duckdb())

# make an in-memory db and store the connection in a variable
duckdb_register(con, "jobs", read_parquet('datasets/jobs.parquet'))
duckdb_register(con, "skills", read_parquet('datasets/skills.parquet'))
duckdb_register(con, "job_skills", read_parquet('datasets/job_skills.parquet'))

```
### SQL example

You can use SQL to query tables. 


```{r sql-ex}

# grouping example using one table
ex1 <- dbGetQuery(con, "SELECT search_position, COUNT(*) AS count_search_position 
                  FROM jobs WHERE job_location = 'Austin, TX' 
                  GROUP BY search_position 
                  ORDER BY count_search_position 
                  DESC LIMIT 10")
print(ex1)


# join example
ex2 <- dbGetQuery(con,"SELECT job_title, skill_name
                  FROM jobs AS j
                  LEFT JOIN job_skills AS x
                    ON x.job_id = j.job_id 
                  LEFT JOIN skills AS s 
                    ON s.skill_id = x.skill_id 
                  LIMIT 10")
print(ex2)

```
### Mixing File Types
You can also mix-and-match file types. Here I am registering a CSV file for the companies tables and joining via SQL to the jobs table.

```{r mix-n-match}
duckdb_register(con, "companies",  read_csv('datasets/companies.csv'))

ex3 <- dbGetQuery(con,"SELECT job_title, company
                  FROM jobs AS j
                  LEFT JOIN companies AS c
                    ON c.company_id = j.company_id
                  LIMIT 10")
print(ex3)                
```


### DuckDB to Dataframe
Using _dbGetQuery_ in the examples above returns a dataframe. Alternately, use the _tbl_ function to access the DuckDB table as a Dataframe. 

```{r to-df }
# transform from sql query
alter_ex3 <- ex3 |>
  mutate(company_2 = substr(company, 2, 5))

glimpse(alter_ex3)


# use tbl to get raw table
jobs <- tbl(con, "jobs")
skills <- tbl(con, "skills")
job_skills <- tbl(con, "job_skills")
companies <- tbl(con, "companies")

jobs <- jobs |>
  mutate(test = 'Hi Team')
glimpse(jobs)

```

More info [here](https://duckdb.org/docs/api/r) 

